Full-day, in home validation of infant body position measurements from inertial sensors

### 2023-03-08 manuscript.Rmd
more kappa

metrics_summary <- metrics_long %>% group_by(Model, metric) %>% 
  get_summary_stats(value) %>% 
  select(-(variable:max), -(q1:mad), -se, -ci) %>% 
  pivot_wider(names_from = Model, values_from = c("median", "mean", "sd")) %>% 
  filter(metric != "Neg Pred Value", metric != "Specificity") %>% 
  relocate(mean_Group, .after = median_Group) %>% 
  relocate(sd_Group, .after = mean_Group) %>% 
  mutate(across(-metric, ~ apa_num(.x, digits = 3)),
         metric = factor(metric, levels = metric_labels)) %>% 
  arrange(metric)

@@

metrics_long_summary <- metrics_long_class %>% group_by(Class,Model, metric) %>% 
  get_summary_stats(value) %>% 
  select(-(variable:max), -(q1:mad), -se, -ci) %>% 
  pivot_wider(names_from = Model, values_from = c("median", "mean", "sd")) %>% 
  filter(metric != "Neg Pred Value", metric != "Specificity") %>% 
  relocate(mean_Group, .after = median_Group) %>% 
  relocate(sd_Group, .after = mean_Group) %>% 
  mutate(across(-(Class:metric), ~ apa_num(.x, digits = 3)),
         metric = factor(metric, levels = metric_labels),
         Class = factor(Class, levels = c("Supine", "Prone", "Sitting", "Upright", "Held"))) %>% 
  relocate(metric, .before = Class) %>% 
  arrange(metric, Class)

@@

apa_table(class_table, col.names = c("Metric","Position", "Median", "Mean", "SD", "Median", "Mean", "SD"), col_spanners = list("Group" = c(3, 5), "Individual" = c(6, 8)),
          midrules = c(5, 10, 15, 20),
          caption = "Model performance metrics for each body position category, shown separately for group and individual models.")
```

@@

@@

Strong overall accuracy can be misleading when class prevalence is unbalanced. For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \@ref(fig:metrics)), and provide classification metrics for each individual body position (Table \@ref(tab:metrics-by-class)) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with median group kappas (`r metrics_lookup["Kappa", "median_Group"]`) somewhat worse compared with median individual kappas (`r metrics_lookup["Kappa", "median_Individual"]`). Guidelines for interpreting kappa statistics [@LandisKoch1977] consider 0.81–1.00 "Almost Perfect,” 0.61–0.80 “Substantial,” 0.41–0.60 “Moderate,” 0.21– 0.40 “Fair,” and 0–0.20 “Slight to Poor”, indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

>>>>

metrics_summary <- metrics_long %>% group_by(Model, metric) %>% 
  get_summary_stats(value) %>% 
  select(-(variable:max), -(q1:mad), -se, -ci) %>% 
  pivot_wider(names_from = Model, values_from = c("median", "mean", "sd")) %>% 
  filter(metric != "Neg Pred Value", metric != "Specificity", metric != "Balanced Accuracy", metric != "F1") %>% 
  relocate(mean_Group, .after = median_Group) %>% 
  relocate(sd_Group, .after = mean_Group) %>% 
  mutate(across(-metric, ~ apa_num(.x, digits = 3)),
         metric = factor(metric, levels = metric_labels)) %>% 
  arrange(metric)

@@

metrics_long_summary <- metrics_long_class %>% group_by(Class,Model, metric) %>% 
  get_summary_stats(value) %>% 
  select(-(variable:max), -(q1:mad), -se, -ci) %>% 
  pivot_wider(names_from = Model, values_from = c("median", "mean", "sd")) %>% 
  filter(metric != "Neg Pred Value", metric != "Specificity", metric != "Balanced Accuracy", metric != "F1") %>% 
  relocate(mean_Group, .after = median_Group) %>% 
  relocate(sd_Group, .after = mean_Group) %>% 
  mutate(across(-(Class:metric), ~ apa_num(.x, digits = 3)),
         metric = factor(metric, levels = metric_labels),
         Class = factor(Class, levels = c("Supine", "Prone", "Sitting", "Upright", "Held"))) %>% 
  relocate(metric, .before = Class) %>% 
  arrange(metric, Class)

@@

apa_table(class_table, col.names = c("Metric","Position", "Median", "Mean", "SD", "Median", "Mean", "SD"), col_spanners = list("Group" = c(3, 5), "Individual" = c(6, 8)),
          midrules = c(5, 10),
          caption = "Model performance metrics for each body position category, shown separately for group and individual models.")
metrics_long_summary <- metrics_long_summary %>% 
  mutate(rowname = paste(Class, metric)) %>% 
  column_to_rownames("rowname")
```

@@

### Overall Accuracy

@@

### Cohen's Kappa 
Strong overall accuracy can be misleading when class prevalence is unbalanced. For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \@ref(fig:metrics)B), and provide classification metrics for each individual body position (Table \@ref(tab:metrics-by-class)) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with median group kappas (`r metrics_lookup["Kappa", "median_Group"]`) somewhat worse compared with median individual kappas (`r metrics_lookup["Kappa", "median_Individual"]`). Guidelines for interpreting kappa statistics [@LandisKoch1977] consider 0.81–1.00 "Almost Perfect,” 0.61–0.80 “Substantial,” 0.41–0.60 “Moderate,” 0.21– 0.40 “Fair,” and 0–0.20 “Slight to Poor”, indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

As in past work, all body positions were accurately classified even though performance varied somewhat between positions. As Table \@ref(tab:metrics-by-class) shows, median kappa statistics were strongest for prone (group = `r metrics_long_summary["Prone Kappa", "median_Group"]`, individual = `r metrics_long_summary["Prone Kappa", "median_Individual"]`) and supine (group = `r metrics_long_summary["Supine Kappa", "median_Group"]`, individual = `r metrics_long_summary["Supine Kappa", "median_Individual"]`). Sitting performance fell in the middle, and was considerably worse for group models than individual models (group = `r metrics_long_summary["Sitting Kappa", "median_Group"]`, individual = `r metrics_long_summary["Sitting Kappa", "median_Individual"]`). Held (group = `r metrics_long_summary["Held Kappa", "median_Group"]`, individual = `r metrics_long_summary["Held Kappa", "median_Individual"]`) and upright (group = `r metrics_long_summary["Upright Kappa", "median_Group"]`, individual = `r metrics_long_summary["Upright Kappa", "median_Individual"]`) performance was the least accurate, however, median performance was still within the "Substantial" range.

### Sensitivity and Positive Predictive Value

### 2023-03-08 manuscript.tex
more kappa

\begin{tabular}{lllllll}
\toprule
 & \multicolumn{3}{c}{Group} & \multicolumn{3}{c}{Individual} \\
\cmidrule(r){2-4} \cmidrule(r){5-7}
Metric & Median & Mean & SD & Median & Mean & SD\\
\midrule
Overall Accuracy & 0.894 & 0.846 & 0.131 & 0.933 & 0.916 & 0.072\\
Balanced Accuracy & 0.913 & 0.891 & 0.077 & 0.911 & 0.907 & 0.071\\
F1 & 0.822 & 0.816 & 0.106 & 0.882 & 0.865 & 0.091\\
Sensitivity & 0.856 & 0.825 & 0.127 & 0.847 & 0.841 & 0.119\\
Pos Pred Value & 0.826 & 0.810 & 0.125 & 0.928 & 0.899 & 0.107\\
Kappa & 0.768 & 0.746 & 0.161 & 0.849 & 0.821 & 0.143\\
\bottomrule
\end{tabular}

@@

\begin{tabular}{llllllll}
\toprule
 &  & \multicolumn{3}{c}{Group} & \multicolumn{3}{c}{Individual} \\
\cmidrule(r){3-5} \cmidrule(r){6-8}
Metric & Position & Median & Mean & SD & Median & Mean & SD\\
\midrule
Balanced Accuracy & Supine & 0.964 & 0.925 & 0.101 & 0.994 & 0.959 & 0.086\\
 & Prone & 0.996 & 0.941 & 0.121 & 0.968 & 0.915 & 0.135\\
 & Sitting & 0.909 & 0.862 & 0.141 & 0.963 & 0.942 & 0.070\\
 & Upright & 0.907 & 0.849 & 0.155 & 0.938 & 0.874 & 0.123\\
 & Held & 0.901 & 0.871 & 0.110 & 0.868 & 0.844 & 0.154\\ \midrule
F1 & Supine & 0.943 & 0.867 & 0.176 & 0.986 & 0.938 & 0.127\\
 & Prone & 0.979 & 0.897 & 0.190 & 0.957 & 0.880 & 0.191\\
 & Sitting & 0.879 & 0.803 & 0.217 & 0.951 & 0.924 & 0.099\\
 & Upright & 0.738 & 0.722 & 0.246 & 0.844 & 0.766 & 0.231\\
 & Held & 0.771 & 0.779 & 0.151 & 0.839 & 0.768 & 0.238\\ \midrule
Sensitivity & Supine & 1.000 & 0.905 & 0.180 & 1.000 & 0.954 & 0.129\\
 & Prone & 1.000 & 0.894 & 0.234 & 0.974 & 0.849 & 0.272\\
 & Sitting & 0.910 & 0.811 & 0.257 & 0.964 & 0.915 & 0.136\\
 & Upright & 0.837 & 0.730 & 0.297 & 0.891 & 0.786 & 0.253\\
 & Held & 0.852 & 0.772 & 0.228 & 0.773 & 0.702 & 0.312\\ \midrule
Pos Pred Value & Supine & 0.995 & 0.828 & 0.292 & 1.000 & 0.932 & 0.134\\
 & Prone & 0.987 & 0.877 & 0.236 & 1.000 & 0.892 & 0.210\\
 & Sitting & 0.896 & 0.802 & 0.261 & 0.972 & 0.945 & 0.079\\
 & Upright & 0.839 & 0.739 & 0.283 & 0.923 & 0.825 & 0.228\\
 & Held & 0.852 & 0.794 & 0.240 & 0.943 & 0.899 & 0.134\\ \midrule
Kappa & Supine & 0.907 & 0.764 & 0.295 & 0.983 & 0.912 & 0.166\\
 & Prone & 0.968 & 0.860 & 0.259 & 0.942 & 0.841 & 0.246\\
 & Sitting & 0.816 & 0.702 & 0.297 & 0.915 & 0.887 & 0.127\\
 & Upright & 0.707 & 0.673 & 0.281 & 0.822 & 0.741 & 0.236\\
 & Held & 0.732 & 0.726 & 0.209 & 0.826 & 0.727 & 0.277\\
\bottomrule
\end{tabular}

@@

Strong overall accuracy can be misleading when class prevalence is unbalanced. For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}), and provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with median group kappas (0.77) somewhat worse compared with median individual kappas (0.85). Guidelines for interpreting kappa statistics (Landis \& Koch, 1977) consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

>>>>

\begin{tabular}{lllllll}
\toprule
 & \multicolumn{3}{c}{Group} & \multicolumn{3}{c}{Individual} \\
\cmidrule(r){2-4} \cmidrule(r){5-7}
Metric & Median & Mean & SD & Median & Mean & SD\\
\midrule
Overall Accuracy & 0.894 & 0.846 & 0.131 & 0.933 & 0.916 & 0.072\\
Kappa & 0.768 & 0.746 & 0.161 & 0.849 & 0.821 & 0.143\\
Sensitivity & 0.856 & 0.825 & 0.127 & 0.847 & 0.841 & 0.119\\
Pos Pred Value & 0.826 & 0.810 & 0.125 & 0.928 & 0.899 & 0.107\\
\bottomrule
\end{tabular}

@@

\begin{tabular}{llllllll}
\toprule
 &  & \multicolumn{3}{c}{Group} & \multicolumn{3}{c}{Individual} \\
\cmidrule(r){3-5} \cmidrule(r){6-8}
Metric & Position & Median & Mean & SD & Median & Mean & SD\\
\midrule
Kappa & Supine & 0.907 & 0.764 & 0.295 & 0.983 & 0.912 & 0.166\\
 & Prone & 0.968 & 0.860 & 0.259 & 0.942 & 0.841 & 0.246\\
 & Sitting & 0.816 & 0.702 & 0.297 & 0.915 & 0.887 & 0.127\\
 & Upright & 0.707 & 0.673 & 0.281 & 0.822 & 0.741 & 0.236\\
 & Held & 0.732 & 0.726 & 0.209 & 0.826 & 0.727 & 0.277\\ \midrule
Sensitivity & Supine & 1.000 & 0.905 & 0.180 & 1.000 & 0.954 & 0.129\\
 & Prone & 1.000 & 0.894 & 0.234 & 0.974 & 0.849 & 0.272\\
 & Sitting & 0.910 & 0.811 & 0.257 & 0.964 & 0.915 & 0.136\\
 & Upright & 0.837 & 0.730 & 0.297 & 0.891 & 0.786 & 0.253\\
 & Held & 0.852 & 0.772 & 0.228 & 0.773 & 0.702 & 0.312\\ \midrule
Pos Pred Value & Supine & 0.995 & 0.828 & 0.292 & 1.000 & 0.932 & 0.134\\
 & Prone & 0.987 & 0.877 & 0.236 & 1.000 & 0.892 & 0.210\\
 & Sitting & 0.896 & 0.802 & 0.261 & 0.972 & 0.945 & 0.079\\
 & Upright & 0.839 & 0.739 & 0.283 & 0.923 & 0.825 & 0.228\\
 & Held & 0.852 & 0.794 & 0.240 & 0.943 & 0.899 & 0.134\\
\bottomrule
\end{tabular}

@@

Strong overall accuracy can be misleading when class prevalence is unbalanced. For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}B), and provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with median group kappas (0.77) somewhat worse compared with median individual kappas (0.85). Guidelines for interpreting kappa statistics (Landis \& Koch, 1977) consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

As in past work, all body positions were accurately classified even though performance varied somewhat between positions. As Table \ref{tab:metrics-by-class} shows, median kappa statistics were strongest for prone (group = 0.968, individual = (0.942)) and supine (group = 0.907, individual = (0.983)). Sitting performance fell in the middle, and was considerably worse for group models than individual models (group = 0.816, individual = (0.915)). Held (group = 0.732, individual = (0.826)) and upright (group = 0.707, individual = (0.822)) performance was the least accurate, however, median performance was still within the ``Substantial'' range.
